Inference for multivariate peaks over threshold models

### 2021-09-16 methodology_pnorm.tex
bruno's edits 2

\begin{figure}[ht]
    \centering
    \includegraphics{images/p_sphere}
    \caption{Unit hyperspheres established on $\mathcal{L}_p$, for $p = 1,\ldots,10$, for two dimensions.\label{fig:psphere}}
\end{figure}

A hypersphere is a geometric object such that the distance from any point to the center takes a fixed,
  constant value.  The unit hypersphere is a hypersphere where that distance is 1. We can define the
  hypersphere under an arbitrary distance measurement, but for our purpose we will use the class of
  hyperspheres defined under the $\mathcal{L}_p$ norm. Let the $\mathcal{L}_p$-norm be defined as
  \begin{equation*}
    \lVert \bm{s} \rVert_p = \left({\textstyle\sum}_{\ell = 1}^d \lvert s_{\ell}\rvert^p\right)^{\frac{1}{p}}.
  \end{equation*}
  From this, we establish the $\mathcal{L}_1$ norm as $p = 1$, or the absolute sum; the
  $\mathcal{L}_2$ norm, as $p = 2$, the Euclidean distance.  From this we also establish the
  $\mathcal{L}_{\infty}$ norm, as
  \begin{equation*}
    \lVert \bm{s} \rVert_{\infty}
      = \lim\limits_{p\to\infty} \lVert \bm{s} \rVert_p
      = \max_{\ell\in\lbrace1,\ldots,d\rbrace}s_{\ell}.
  \end{equation*}
  We are interested in the direction, or angular component, of vectors described in the positive
  orthant, $\mathcal{R}_{+}^d$.  One means of describing the direction of a vector in $\mathcal{R}_+^d$
  is to project that vector onto $\mathcal{S}_{p}^{d-1}$, the positive orthant of the unit hypersphere
  defined on an $\mathcal{L}_p$-norm, denoted as $\mathcal{S}_{p}^{d-1}$.  That is,
  \begin{equation*}
    \mathcal{S}_{p}^{d-1} = \left\lbrace \bm{s} : \bm{s} \in \mathcal{R}_{+}^{d}, \lVert \bm{s}\rVert_{p} = 1\right\rbrace.
  \end{equation*}
  Figure~\ref{fig:psphere} shows $\mathcal{S}_{p}^{1}$ for $p$ at integers between 1 and 10.
  At $p = 1$, we have the simplex.  As $p$ grows larger, the surface moves away from the origin and
  approaches closer to the unit hypercube--the limiting form of the surface. We project an observation
  onto $\mathcal{S}_p^{d-1}$ by dividing said observation by its $p$-norm--let
  $\bm{x}\in \mathcal{R}_{+}^{d}$, then $\bm{y} = \bm{x} / \lVert \bm{x}\rVert_p \in \mathcal{S}_{p}^{d-1}$.
  We denote the $d-1$ to indicate the loss of one degree of freedom relative to the original vector.
  Observations on one hypersphere can be projected without loss of information onto another by
  dividing by the defining norm of the target hypersphere.

Assuming $\bm{y} \in \mathcal{S}_{p}^{d-1}$, then for finite $p$, $y_d$ can always be represented
  as a function of the other dimensions.  That is,
  \begin{equation*}
    y_d = \left(1 - {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p\right)^{\frac{1}{p}}.
  \end{equation*}
  So the transformation
  \begin{equation}
    \label{eqn:pnormt}
    T(x_1,\ldots,x_d) = \left(\pnorm{\bm{x}}{p}, \frac{x_1}{\pnorm{\bm{x}}{p}},
                          \ldots , \frac{x_{d-1}}{\pnorm{\bm{x}}{p}}\right) = (r,y_1,\ldots,y_{d-1})
  \end{equation}
  does not lose any information.  The reverse of this transformation,
  \begin{equation}
    \label{eqn:pnormtinv}
    T^{-1}\left(r,y_1,\ldots,y_{d-1}\right) =
      \left(ry_1,\ldots,ry_{d-1}, r\left(1 - {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p\right)^{\frac{1}{p}}\right)
  \end{equation}
  equivalently recovers the original data.  To recover the density of the transformed random variables,
  the original density is multiplied by the determinant of the Jacobian--the matrix of derivatives
  of the inverse transformation.  This takes the form:
  \begin{equation}
    \label{eqn:pnormjac}
    r^{d-1}\left[\left(1 - {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p\right)^{\frac{1}{p}} +
        {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(1 - {\textstyle\sum}_{l=1}^{d-1} y_{\ell}^p\right)^{\frac{1}{p} - 1}\right].
  \end{equation}
  Notice a factor of $r^{d-1}$ independent of $p$. We refer to $\bm{y}$ and $r$ as, respectively,
  the angular and radial components of $\bm{x}$.  If we assume a distribution for $\bm{x}$, then
  by transforming to $r, \bm{y}$ and integrating out $r$, we are left with a distribution on solely
  $\bm{y}$, the projection of $\bm{x}$ onto $\mathcal{S}_{p}^{d-1}$.  If we take the projection to
  its limit as $p\to\infty$, we arrive at a projection onto $\mathcal{S}_{\infty}^{d-1}$.

We can observe a problem at this point in at least two ways.  First, on $\mathcal{S}_{\infty}^{d-1}$, the
  inverse transformation from $\mathcal{S}_{\infty}^{d-1}$ to $R_+^{d}$ in Equation~\ref{eqn:pnormtinv}
  may not recover the original data--that formulation of the projection would only work when $Y_d = 1$.
  Alternatively, we may look at the formula for the determinant of the Jacobian in
  Equation~\ref{eqn:pnormjac}.  In that formula we observe the value
  $(1 - \sum_{\ell = 1}^{d-1}y_{\ell}^p)^{\frac{1}{p} - 1}$.  As $p\to\infty$,
  this quantity will go to $0^{-1}$ if $y_d \neq 1$.
  So, the mathematical basis of this transformation is thus dependent upon finite $p$.  However, we can
  project onto $\mathcal{S}_p^{d-1}$ for large $p$, and that projection will be similar to an idealized
  distribution on $\mathcal{S}_{\infty}^{d-1}$.

>>>>

%\begin{figure}[ht]
%    \centering
%    \includegraphics{images/p_sphere}
%    \caption{Unit hyperspheres established on $\mathcal{L}_p$, for $p = 1,\ldots,10$, for two dimensions.\label{fig:psphere}}
%\end{figure}

Consider a vector $\bm{s} \in {\mathbb R}^d$, and let the $\mathcal{L}_p$-norm be defined as
  \begin{equation*}
    \lVert \bm{s} \rVert_p = \left({\textstyle\sum}_{\ell = 1}^d \lvert s_{\ell}\rvert^p\right)^{\frac{1}{p}},
  \end{equation*}
for $p>0$. The absolute and Euclidean norms are obtained for $p=1$ and $p=2$ respectively. 
The $\mathcal{L}_{\infty}$ norm can be obtained as a limit, 
  \begin{equation*}
    \lVert \bm{s} \rVert_{\infty}
      = \lim\limits_{p\to\infty} \lVert \bm{s} \rVert_p
      = \max_{\ell\in\lbrace1,\ldots,d\rbrace}s_{\ell}.
  \end{equation*}
To define an angular measure,  we are interested in the direction of
vectors in ${\mathbb R}_{+}^d$, thus, we project them onto
  \begin{equation*}
    \mathcal{S}_{p}^{d-1} = \left\lbrace \bm{s} : \bm{s} \in {\mathbb R}_{+}^{d}, \lVert \bm{s}\rVert_{p} = 1\right\rbrace,
  \end{equation*}
  the $\mathcal{L}_p$-norm unit hypersphere.
  Given $\bm{x}\in {\mathbb R}^d_+$ the projection onto $\mathcal{S}_{p}^{d-1}$ is given as
  $\bm{y} = \bm{x} / \lVert \bm{x}\rVert_p \in \mathcal{S}_{p}^{d-1}$.
  Observations on one hypersphere can be projected without loss of information onto another by
  dividing by the defining norm of the target hypersphere. In fact, letting 
  \begin{equation*}
    y_d = \left(1 - {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p\right)^{\frac{1}{p}}.
  \end{equation*}
  the transformation
  \begin{equation}
    \label{eqn:pnormt}
    T(x_1,\ldots,x_d) = \left(\pnorm{\bm{x}}{p}, \frac{x_1}{\pnorm{\bm{x}}{p}},
                          \ldots , \frac{x_{d-1}}{\pnorm{\bm{x}}{p}}\right) = (r,y_1,\ldots,y_{d-1})
  \end{equation}
  is invertible with
  \begin{equation}
    \label{eqn:pnormtinv}
    T^{-1}\left(r,y_1,\ldots,y_{d-1}\right) =
      \left(ry_1,\ldots,ry_{d-1}, r\left(1 - {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p\right)^{\frac{1}{p}}\right).
  \end{equation}
  The Jacobian of the transformation takes the form
  \begin{equation}
    \label{eqn:pnormjac}
    r^{d-1}\left[\left(1 - {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p\right)^{\frac{1}{p}} +
        {\textstyle\sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(1 - {\textstyle\sum}_{l=1}^{d-1} y_{\ell}^p\right)^{\frac{1}{p} - 1}\right].
  \end{equation}
  Equations (\ref{eqn:pnormt}) -- (\ref{eqn:pnormjac}) can be used to define a density on ${\mathbb R}^p_+$, and then obtain a density on $\mathcal{S}_{p}^{d-1}$, for any finite $p>0$. Unfortunately in the limit, when $p\rightarrow +\infty$, the transformation is not differentiable. Thus, even though the projected distribution on $\mathcal{S}_\infty^{d-1}$ is defined, there is no corresponding density.

### 2021-09-16 methodology_projgamma.tex
bruno's edits 2

\subsection{Projected Gamma Family}
As stated previously, one approach we might try in order to generate a distribution on
  $\mathcal{S}_{p}^{d-1}$ would be to take a distribution in $\mathcal{R}_{+}^d$, reparameterize
  it in terms of radial and angular components, and integrate out the radial component in order to
  effectively project the distribution onto $\mathcal{S}_{p}^{d-1}$.  A natural candidate distribution
  to accomplish this is the Gamma.  Let
    $\bm{ X} \sim \prod_{\ell = 1}^d\text{Ga}\left(X_{\ell}\mid\alpha_{\ell},\beta_{\ell}\right)$.
  Then transform $\bm{ X}$ via the transformation in Eqns.~\ref{eqn:pnormt} into $R,\bm{ Y}$, where
  $R = \pnorm{\bm{ X}}{p}$ and $Y\in \mathcal{S}_{p}^{d-1}$.  Completing the transformation, we
  incorporate the determinant of the Jacobian from Eqn.~\ref{eqn:pnormjac}.  Thus we have
  \begin{equation}
    \label{eqn:protopg}
    f(r,\bm{ y}) = \prod_{\ell = 1}^{d}
      \left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}(ry_{\ell})^{\alpha_{\ell} - 1}
          \exp\lbrace-\beta_{\ell}ry_{\ell}\rbrace\right]
      \times r^{d-1}\left[y_d -
            {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right],
  \end{equation}
  where $y_d = \left(1 - \sum_{\ell = 1}^{d-1}y_{\ell}^p\right)^{\frac{1}{p}}$. Integrating out $r$
  yields the resulting \emph{Projected Gamma} density
  \begin{equation}
    \label{eqn:projgamma}
    \text{PG}(\bm{ y}\mid\bm{ \alpha},\bm{ \beta}) =
          \prod_{\ell = 1}^d\left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}
                y_{\ell}^{\alpha_{\ell} - 1}\right]
      \times \left[y_d -
          {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right]
      \times \frac{\Gamma({\textstyle\sum}_{\ell = 1}^d\alpha_{\ell})}{\left({\textstyle\sum}_{\ell = 1}^d
                    \beta_{\ell}y_{\ell}\right)^{{\scriptstyle\sum_{\ell = 1}^d \alpha_{\ell}}}}
  \end{equation}
  established for an arbitrary $\mathcal{L}_p$ norm. This forms a valid density for all finite $p$.
  We can regard the Dirichlet distribution as a special case of this density, where
  $\bm{ y}\in \mathcal{S}_1^{d-1}$, and $\beta_{\ell} = \beta$ for all $\ell$.  The distribution
  described in \cite{nunez2019} is related to this distribution in that it is established on
  $\bm{ y}\in \mathcal{S}_2^{d-1}$, then transformed to spherical coordinates via Eqn.~\ref{eqn:spherical}.
  The full conditional for $r$ takes the form
  \begin{equation}
    \label{eqn:rfullcond}
    r\mid\bm{ \alpha},\bm{ \beta}, y \sim \text{Ga}\left(r\mid{\textstyle\sum}_{\ell = 1}^d
                    \alpha_{\ell}, {\textstyle\sum}_{\ell = 1}^d \beta_{\ell}y_{\ell}\right).
  \end{equation}

In the estimation of the parameters of this model, we can quickly arrive at a problem, in that
  without some linear restriction on the rate parameters $\beta_{\ell}$, the model is not identifiable.
  The restriction most often used in the case of the Generalized Dirichlet, as well as by
  \cite{nunez2019} is to set $\beta_1 := 1$.

\subsubsection{Inference on Projected Gamma Model}
We consider a basic model that assumes all observations are generated by a single projected Gamma
  distribution.  The shape and rate parameters for the latent \text{Gamma} random variables,
  $\alpha_{\ell},\beta_{\ell}$, must be greater than 0, so a logical prior to use would be a Gamma prior for
  each parameter, excepting for $\beta_1 := 1$.
  \begin{equation}
    \label{eqn:vanillap}
    \begin{aligned}
      \bm{ y}\mid\alpha,\beta &\sim \text{PG}(\bm{ y}\mid\alpha,\beta)\\
      \bm{ \alpha},\bm{\beta} &\sim {\textstyle \prod}_{\ell = 1}^d \text{Ga}(\alpha_{\ell} \mid \xi_{\ell},\tau_{\ell})
              \times {\textstyle \prod}_{\ell = 2}^d \text{Ga}(\beta_{\ell}\mid \zeta_{\ell},\sigma_{\ell}).
    \end{aligned}
  \end{equation}
  A useful property of the Projected Gamma family is the latent independent Gammas assumption.  Recall
  in Eqn.~\ref{eqn:protopg}, we have a density defined as a product of independent Gamma densities,
  multiplied by a factor of $r^{d-1}C$, where $C$ is constant with respect to the parameters
  $\alpha,\beta$, as well as $r$.  If we conduct inference via a Gibbs sampler, then by simulating
  this latent $r$ we are able to perform inference of $(\alpha_{\ell},\beta_{\ell})$ independent of
  that of $(\alpha_k,\beta_k)$.  The full conditional for $r$ is described in Eqn.~\ref{eqn:rfullcond}.
  The full conditional for $\beta_{\ell}$, $l = 2,\ldots,d$ then becomes
  \begin{equation}
    \label{eqn:betafc}
    \beta_{\ell}\mid \bm{ y}, r, \alpha_{\ell} \sim \text{Ga}\left(n + \alpha_{\ell} + \zeta_{\ell},
                                      {\textstyle \sum}_{i = 1}^nr_iy_{il} + \sigma_{\ell}\right).
  \end{equation}
  That is, given $\alpha_{\ell}$, $r$, we can sample $\beta_{\ell}$ directly as a Gamma random
  variable.  In the sampling for $\alpha_{\ell}$, we can integrate $\beta_{\ell}$ out leaving us a
  quantity proportional to a density, which we can sample via the Metropolis Hastings
  algorithm\citep{hastings1970}.  For $\alpha_{\ell}$, $\ell = 2,\ldots, d$, This takes the form
  \begin{equation}
    \label{eqn:alphafc}
    f(\alpha_{\ell} \mid \bm{ y}, r) \propto
    \frac{\left({\textstyle \prod}_{i = 1}^nr_iy_{il}\right)^{\alpha_{\ell} - 1}}{\Gamma^n(\alpha_{\ell})} \times
    \alpha_{\ell}^{\xi_{\ell} - 1}\exp\{-\tau_{\ell}\alpha_{\ell}\} \times
    \frac{\Gamma(n\alpha_{\ell} + \zeta_{\ell})}{
        \left({\textstyle\sum}_{i = 1}^n r_iy_{il} + \sigma_{\ell}\right)^{(n * \alpha_{\ell} + \zeta_{\ell})}}.
  \end{equation}
  For $\alpha_1$, as we do not integrate out $\beta_1$ as it is defined to be $1$, this takes the
  slightly simpler form
  \begin{equation}
    \label{eqn:alphafc1}
    f(\alpha_1 \mid \bm{ y}, r) \propto
      \frac{{\textstyle\prod}_{i = 1}^n ry_{i1}^{\alpha_1 - 1}}{\Gamma^n(\alpha_1)} \times
      \alpha_1^{\xi_1 - 1}\exp\{-\tau_1\alpha_1\}.
  \end{equation}
  The \emph{projected restricted Gamma} model sets $\beta_{\ell} := 1 \forall \ell \in \lbrace 1,\ldots,d\rbrace$,
  so in that model inference on all $\alpha_{\ell}$ will use the simpler form.

>>>>

\subsection{Projected gamma family}
  A natural  distribution to consider in ${\mathbb R}^p_+$ is given by a product of independent
  univariate Gamma distributions. Let
    $\bm{ X} \sim \prod_{\ell = 1}^d\text{Ga}\left(X_{\ell}\mid\alpha_{\ell},\beta_{\ell}\right)$, 
    $a_\ell$ and $b_\ell$ are the shape and scale parameters, repectively. Using Equations (\ref{eqn:pnormt}) -- (\ref{eqn:pnormjac}) we have the joint density
  \begin{equation*}
    f(r,\bm{ y}) = \prod_{\ell = 1}^{d}
      \left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}(ry_{\ell})^{\alpha_{\ell} - 1}
          \exp\lbrace-\beta_{\ell}ry_{\ell}\rbrace\right]
      \times r^{d-1}\left[y_d -
            {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right],
  \end{equation*}
  Integrating out $r$ yields the resulting \emph{Projected Gamma} density
  \begin{equation}
    \label{eqn:projgamma}
    \text{PG}(\bm{ y}\mid\bm{ \alpha},\bm{ \beta}) =
          \prod_{\ell = 1}^d\left[\frac{\beta_{\ell}^{\alpha_{\ell}}}{\Gamma(\alpha_{\ell})}
                y_{\ell}^{\alpha_{\ell} - 1}\right]
      \times \left[y_d -
          {\textstyle \sum}_{\ell = 1}^{d-1}y_{\ell}^p\left(y_d^p\right)^{\frac{1}{p} - 1}\right]
      \times \frac{\Gamma({\textstyle\sum}_{\ell = 1}^d\alpha_{\ell})}{\left({\textstyle\sum}_{\ell = 1}^d
                    \beta_{\ell}y_{\ell}\right)^{{\scriptstyle\sum_{\ell = 1}^d \alpha_{\ell}}}}
  \end{equation}
  defined for $\bm{y}\in \mathcal{S}_p^{d-1}$, and for any $p>0$. To avoid idetifiability problems when estimating the shape and scale parameters, we set $\beta_1 = 1$.
  \cite{nunez2019} obtain the density in Equation (\ref{eqn:projgamma})
  for $p=2$ as a multivariate distribution for directional data, using spherical coordinates.
  For $\bm{ y}\in \mathcal{S}_1^{d-1}$, and $\beta_{\ell} = \beta, \forall\ell$, the density in Equation (\ref{eqn:projgamma}) corresponds to that of a Dirichlet distribution.

  The projected gamma family is simple to specify and has very tractable computational properties. Thus, we use it as a building block for the angular measure $\Phi$ models. To build a flexible family of distributions in $\mathcal{S}_p^{d-1}$ we consider mixtures of projected gamma densities defined as
  \begin{equation} \label{PGmix}     
     f(\bm{y}) = \int_\Theta \text{PG}(\bm{y}|\bm{\theta}) dG(\bm{\theta}), 
  \end{equation} 
  where $\bm{\theta} = (\bm{\alpha}, \bm{\beta})$. Following a Bayesian non-parametric approach \citep{Ferguson74,Antoniak1974} we assume that $G$ is drawn from a random measure. In particular, assuming a Dirichlet process prior, we have a hierarchical formulation of the mixture model given by
  \[    \bm{y} \sim \text{PG}(\bm{y}|\bm{\theta}) , \;\;\; \bm{\theta} \sim G, \;\;\; G \sim \text{DP}(\eta, G_0),\]
  where DP denotes a Dirichlet process, $\eta$ is the precision parameter, and $G_0$ is the centering distribution.

  Our strategy consists of describing the angular distribution $\Phi$ using a sample based approach with the following steps: (i) Apply the transformation in Equation (\ref{eqn:standardization}) to the original data; (ii) Obtain the subsample of the standardized observations that satisfy $R>1$; (iii) Take a finite $p$ and project the observations onto $\mathcal{S}_p^{d-1}$; (iv) Fit the model in Equation (\ref{PGmix}) to the resulting data and obtain samples from the fitted model; (v) Project the resulting samples to $\mathcal{S}_\infty^{d-1}$. 
  {\bf Instead of the olf Figure 1, it would be nice to add a figure where you illustrate this procedure. Give it a try.}
%  The full conditional for $r$ takes the form
%  \begin{equation}
%    \label{eqn:rfullcond}
%    r\mid\bm{ \alpha},\bm{ \beta}, y \sim \text{Ga}\left(r\mid{\textstyle\sum}_{\ell = 1}^d
%                    \alpha_{\ell}, {\textstyle\sum}_{\ell = 1}^d \beta_{\ell}y_{\ell}\right).
%  \end{equation}

\subsubsection{Inference for the projected gamma mixture model}
{\bf Our focus is on model (\ref{PGmix}). Describe how you build a sample based approach to fitting that model. How you introduce latent variables, both to break the mixture and to produce the PG. You can then discuss variations that essentially consists on different priors. So, basically, you need to condense the next three pages in about a page.}

We consider a basic model that assumes all observations are generated by a single projected Gamma
  distribution.  The shape and rate parameters for the latent \text{Gamma} random variables,
  $\alpha_{\ell},\beta_{\ell}$, must be greater than 0, so a logical prior to use would be a Gamma prior for
  each parameter, excepting for $\beta_1 := 1$.
  \begin{equation}
    \label{eqn:vanillap}
    \begin{aligned}
      \bm{ y}\mid\alpha,\beta &\sim \text{PG}(\bm{ y}\mid\alpha,\beta)\\
      \bm{ \alpha},\bm{\beta} &\sim {\textstyle \prod}_{\ell = 1}^d \text{Ga}(\alpha_{\ell} \mid \xi_{\ell},\tau_{\ell})
              \times {\textstyle \prod}_{\ell = 2}^d \text{Ga}(\beta_{\ell}\mid \zeta_{\ell},\sigma_{\ell}).
    \end{aligned}
  \end{equation}
  A useful property of the Projected Gamma family is the latent independent Gammas assumption.  Recall
  in Eqn.~\ref{eqn:protopg}, we have a density defined as a product of independent Gamma densities,
  multiplied by a factor of $r^{d-1}C$, where $C$ is constant with respect to the parameters
  $\alpha,\beta$, as well as $r$.  If we conduct inference via a Gibbs sampler, then by simulating
  this latent $r$ we are able to perform inference of $(\alpha_{\ell},\beta_{\ell})$ independent of
  that of $(\alpha_k,\beta_k)$.  The full conditional for $r$ is described in Eqn.~\ref{eqn:rfullcond}.
  The full conditional for $\beta_{\ell}$, $l = 2,\ldots,d$ then becomes
  \begin{equation}
    \label{eqn:betafc}
    \beta_{\ell}\mid \bm{ y}, r, \alpha_{\ell} \sim \text{Ga}\left(n + \alpha_{\ell} + \zeta_{\ell},
                                      {\textstyle \sum}_{i = 1}^nr_iy_{il} + \sigma_{\ell}\right).
  \end{equation}
  That is, given $\alpha_{\ell}$, $r$, we can sample $\beta_{\ell}$ directly as a Gamma random
  variable.  In the sampling for $\alpha_{\ell}$, we can integrate $\beta_{\ell}$ out leaving us a
  quantity proportional to a density, which we can sample via the Metropolis Hastings
  algorithm\citep{hastings1970}.  For $\alpha_{\ell}$, $\ell = 2,\ldots, d$, This takes the form
  \begin{equation}
    \label{eqn:alphafc}
    f(\alpha_{\ell} \mid \bm{ y}, r) \propto
    \frac{\left({\textstyle \prod}_{i = 1}^nr_iy_{il}\right)^{\alpha_{\ell} - 1}}{\Gamma^n(\alpha_{\ell})} \times
    \alpha_{\ell}^{\xi_{\ell} - 1}\exp\{-\tau_{\ell}\alpha_{\ell}\} \times
    \frac{\Gamma(n\alpha_{\ell} + \zeta_{\ell})}{
        \left({\textstyle\sum}_{i = 1}^n r_iy_{il} + \sigma_{\ell}\right)^{(n * \alpha_{\ell} + \zeta_{\ell})}}.
  \end{equation}
  For $\alpha_1$, as we do not integrate out $\beta_1$ as it is defined to be $1$, this takes the
  slightly simpler form
  \begin{equation}
    \label{eqn:alphafc1}
    f(\alpha_1 \mid \bm{ y}, r) \propto
      \frac{{\textstyle\prod}_{i = 1}^n ry_{i1}^{\alpha_1 - 1}}{\Gamma^n(\alpha_1)} \times
      \alpha_1^{\xi_1 - 1}\exp\{-\tau_1\alpha_1\}.
  \end{equation}
  The \emph{projected restricted Gamma} model sets $\beta_{\ell} := 1 \forall \ell \in \lbrace 1,\ldots,d\rbrace$,
  so in that model inference on all $\alpha_{\ell}$ will use the simpler form.

### 2021-09-16 results.tex
bruno's edits 2

\section{Results}

\subsection{Simulation}
We have introduced this projected gamma distribution existing on $\mathcal{S}_{\infty}^{p-1}$.  As a
  means of evaluating how well we can recover a distribution under the proposed model, we have
  generated a series of datasets on $\mathcal{S}_{\infty}^{d-1}$, for $d$ ranging between 3 and 20.
  We generated each dataset from a mixture of projected Gammas, with the number of mixture components
  ranging between 3 and 12.  The generation procedure is detailed in Algorithm~\ref{algo:simulated}.
  \begin{algorithm}[htb]
    \For{$n_{\text{mix}}$ in $[3, 6, 9, 12]$}{
      Generate $n_{\text{mix}} \times 20$ shape Parameters $\alpha$\\
      Generate $n_{\text{mix}} \times 20$ Rate Parameters $\beta$\\
      Generate $500$ Mixture Component Identifiers $\delta$\\
      \For{$i$ in $1,\ldots,500$}{
        Generate ${\bf X}_i \sim \prod_{\ell = 1}^d\text{Ga}\left(X_{il}\mid\alpha_{\delta_i,l},\beta_{\delta_i, l}\right)$
        }
      \For{$n_{\text{col}}$ in $[3,6,12,20]$}{
        Project columns 1 to $n_{\text{col}}$ of ${\bf X}$ onto $\mathcal{S}_{\infty}^{n_{\text{col}} - 1}$ and save.
        }
    }
    \caption{Simulated Angular Dataset Generation Routine\label{algo:simulated}}
  \end{algorithm}
  The choice of generation routine for $\alpha = \alpha_0 + \alpha_1$ where
  $\alpha_0 \sim \text{Unif}(\alpha_0\mid 0,4)$, $\alpha_1\sim \text{Gamma}(\alpha_1\mid 1,1)$.
  The choice of generation routine for $\beta$ is $\beta\sim\text{Unif}(\beta\mid 0.25, 2.5)$.
  This choice was made to ensure that $\beta$ does not approach 0 in the generation routine.

@@

\subsection{Integrated Vapor Transport}
We use data from the integrated water vapor transport (\emph{IVT}) model \citep{guan2015}
  of atmospheric rivers as a means of testing these models.  An atmospheric river is a meteorological
  event of local concentration of water vapor in the atmosphere that moves with wind patterns.
  Understanding extremal dependence in atmospheric rivers, for instance, grants hydrological engineers
  a greater understanding of how and where to build ground water capacity to handle precipitation.
  California suffers persistent, extended droughts---if ground water transportation and holding capacity
  (canals, reservoirs) are not built to manage extreme precipitation without adequate understanding of
  the extremal dependence, then we may not install adequate capacity and thus experience flooding.
  Alternatively, if we install too much capacity everywhere, then we are wasting money that could be
  better spent elsewhere.

>>>>

\section{Data illustrations}

\subsection{Simulations}
We have introduced this projected gamma distribution existing on $\mathcal{S}_{\infty}^{p-1}$.  As a
  means of evaluating how well we can recover a distribution under the proposed model, we have
  generated a series of datasets on $\mathcal{S}_{\infty}^{d-1}$, for $d$ ranging between 3 and 20.
  We generated each dataset from a mixture of projected Gammas, with the number of mixture components
  ranging between 3 and 12.  The generation procedure is detailed in Algorithm~\ref{algo:simulated}.
  \begin{algorithm}[htb]
    \For{$n_{\text{mix}}$ in $[3, 6, 9, 12]$}{
      Generate $n_{\text{mix}} \times 20$ shape Parameters $\alpha$\\
      Generate $n_{\text{mix}} \times 20$ Rate Parameters $\beta$\\
      Generate $500$ Mixture Component Identifiers $\delta$\\
      \For{$i$ in $1,\ldots,500$}{
        Generate ${\bf X}_i \sim \prod_{\ell = 1}^d\text{Ga}\left(X_{il}\mid\alpha_{\delta_i,l},\beta_{\delta_i, l}\right)$
        }
      \For{$n_{\text{col}}$ in $[3,6,12,20]$}{
        Project columns 1 to $n_{\text{col}}$ of ${\bf X}$ onto $\mathcal{S}_{\infty}^{n_{\text{col}} - 1}$ and save.
        }
    }
    \caption{Simulated Angular Dataset Generation Routine\label{algo:simulated}}
  \end{algorithm}
  The choice of generation routine for $\alpha = \alpha_0 + \alpha_1$ where
  $\alpha_0 \sim \text{Unif}(\alpha_0\mid 0,4)$, $\alpha_1\sim \text{Gamma}(\alpha_1\mid 1,1)$.
  The choice of generation routine for $\beta$ is $\beta\sim\text{Unif}(\beta\mid 0.25, 2.5)$.
  This choice was made to ensure that $\beta$ does not approach 0 in the generation routine.

@@

\subsection{Integrated Vapor Transport}
One metric by which we might identify and declare atmospheric rivers is the integrated water vapor
  transport, or \emph{IVT}.  This value represents the total amount of water
  vapor being transported in an atmospheric column---that is, a column of the troposphere above a given
  area of the surface of the Earth. Raw data are measured by dropsondes---measurement devices dropped
  out of aircraft, taking measurements as they fall to the earth.~\citep{ralph2017}  A data vector
  includes an estimate of IVT at grid cell at a period in time; where a grid cell specifies the surface
  of the earth under the atmospheric column.  We have two such datasets, in differing spatial
  resolutions---the lower resolution data splits the coast of california into 8 grid cells, while the
  higher resolution does so into 47 grid cells~\citep{guan2015}.

As IVT tracks water vapor in the atmosphere, extreme values in IVT can lead to
  extreme precipitation events.  Atmospheric rivers are necessary events, in that they carry the
  bulk of precipitation that California recieves, but extreme precipitation events can be extremely
  destructive as well.  Thus, we are interested in the distribution of extreme precipitation events,
  including their depenedence structure.  This tells us about the probability of jointly extreme
  events---where IVT is extreme in two or more grid cells, as well as identifying particular
  configurations of IVT that may be anomalous.  In California, groundwater is a managed resource.
  Extreme precipitation events significantly affect groundwater, so anomalous precipitation events
  can produce anomalous configurations of groundwater, leading to destructive flooding.

We use data from the integrated water vapor transport (\emph{IVT}) model \citep{guan2015}
  of atmospheric rivers as a means of testing these models.  An atmospheric river is a meteorological
  event of local concentration of water vapor in the atmosphere that moves with wind patterns.
  Understanding extremal dependence in atmospheric rivers, for instance, grants hydrological engineers
  a greater understanding of how and where to build ground water capacity to handle precipitation.
  California suffers persistent, extended droughts---if ground water transportation and holding capacity
  (canals, reservoirs) are not built to manage extreme precipitation without adequate understanding of
  the extremal dependence, then we may not install adequate capacity and thus experience flooding.
  Alternatively, if we install too much capacity everywhere, then we are wasting money that could be
  better spent elsewhere.

  {\bf We need a one paragraph summary of the stuff in the last three paragraphs. Say what IVT, how it is obtained, why it is important and describe the two datasets you have.}
